### 2018-03-01 paper-amia-2018/manuscript-template.tex
update introduction

\documentclass{amia}
\usepackage{graphicx}
\usepackage[labelfont=bf]{caption}
\usepackage[superscript,nomove]{cite}
\usepackage{color}
\usepackage{multirow}
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

\begin{document}

\title{Summarizes the Hypothesis of the Paper}

\author{Mehedi Hasan, BS$^{1}$\footnote[1]{Authors provided equal contribution. \label{footnote1}}, Alexander Kotov, PhD$^{1}$\textsuperscript{\ref{footnote1}}, April Idalski Carcone, PhD$^{2}$, Ming Dong, PhD$^{1}$, Sylvie Naar, PhD$^{2}$}

\institutes{
$^1$Department of Computer Science, Wayne State University, Detroit, Michigan \\  
$^2$Department of Family Medicine and Public Health Sciences, School of Medicine, Wayne State University, Detroit, Michigan\\
}

\maketitle

\noindent{\bf Abstract}
\textit{\\
1) Motivation: Why do we care about the problem and the results? \\
2) Problem statement: What problem is the paper trying to solve and what is the scope of the work? \\
3) Approach: What was done to solve the problem? \\
4) Results: What is the answer to the problem? \\
5) Conclusions: What implications does the answer imply?}

\section*{Introduction}

What is the problem? \\
Why is the problem important? \\
What has so far been done on the  problem?\\
1) describes previous work in more technical detail \\
2) as far as needed for a proper understanding of the contribution of the paper \\
What is the contribution of the paper on the problem?\\
1) What are the rival approaches? \\
2) What are the drawbacks of each?\\
3) How has the battle between different approaches progressed?\\
4) What are the major outstanding problems? (This is where you come in)\\
Is the contribution original? Explain why \\
Is the contribution non-trivial? Explain why \\
The authors are not aware of any other work where this approach has been considered.

\section*{Methods}
Part 1

1) What algorithms or data structures did you select? Who created them? What is their asymptotic behavior? What other specific characteristics are worth noting for this study?\\
2) What programming language and platform did you use? What impact did these choices have on your project?\\

Part 2

1) How specifically did you implement the algorithms?\\
2) How did you handle instrumentation code? Why?\\
3) Did you perform any optimizations? Why or why not?\\
4) How did you choose to test and benchmark your code?\\
5) What inputs (data) did you select to test your implementations? Why?\\

\section*{Results}

1) In general, the pure, unbiased results should be presented first without interpretation (van Wagenen 1990). \\
2) These results should present the raw data
or the results after applying the techniques outlined in the methods section. The results are simply results; they do not draw conclusions.

\section*{Discussion}
1) What, specifically, did you learn from comparing these algorithms or data structures? \\
2) What do your results say about the problem or question you were investigating?\\
3) Was your hypothesis confirmed or disproved?\\
4) Are the results what you expected?\\
5) If you obtained anomalies or other unexpected results, can you explain them? If not, how could you set about in the future to identify what caused them? \\
6) How do your results compare to past findings? Are they consistent? Different?Why? \\
7) How would you respond to objections or questions that other researchers might have about your methods, results, or interpretations? \\
8) What is new and significant?\\

\section*{Conclusion}
1) The hypothesis and the evidence for and against it are briefly restated \\
2) The original motivation is recapitulated \\
3) The state of the field in the light of this new contribution is reassessed \\
4) describes future research and new directions suggested by the contribution\\
5) in particular, research that would improve the evidence for or against the hypothesis

\section*{Acknowledgments}
Comming soon!

\bibliographystyle{vancouver}
\bibliography{References}

\end{document}